SeedLing: Building and using a seed corpus \\ for the Human Language Project

### 2014-03-20 docs/computEL/cluster.tex
added bib citations to cluster section

When working on low-resource or endangered languages, computational and documentary linguists face the issue of lack of resources and knowledge about the language. Often, having knowledge about related or similar language(s) provides useful lexical, syntactic or morphological minimal pairs across languages and also helps in bootstrapping language models in NLP (Yarowsky
and Ngai, 2001; Xia and Lewis, 2007).

The task of identifying related/similar language is also known as ``language classification" where (i) languages are mapped onto language families trees depending on their historical ancestry often measured by comparing string distance from a fix lexical list across languages, aka. \emph{genetic classification} or \emph{lexicostatistics} (e.g. Swadesh, 1952; Starostin, 2010) or (ii) languages are grouped by their degrees of similarity with regards to their typological features, aka. \emph{typological classification} (e.g. Daume, 2009; Georgi et al. 2010).

>>>>

When working on low-resource or endangered languages, computational and documentary linguists face the issue of lack of resources and knowledge about the language. Often, having knowledge about related or similar language(s) provides useful lexical, syntactic or morphological minimal pairs across languages and also helps in bootstrapping language models in NLP \cite{yarowsky:ngai:2001,xia2007multilingual}.

The task of identifying related/similar language is also known as ``language classification" where (i) languages are mapped onto language families trees depending on their historical ancestry often measured by comparing string distance from a fix lexical list across languages, aka. \emph{genetic classification} or \emph{lexicostatistics} \cite{swadesh1952,starostin2010} or (ii) languages are grouped by their degrees of similarity with regards to their typological features, aka. \emph{typological classification} \cite{georgi2010wals,daume2009}.

### 2014-03-20 docs/computEL/data.tex
various small mods
added bib citations to cluster section

\begin{figure*}[t]
\begin{centering}
\includegraphics[scale=0.6]{heatmap-numbers.png}

\caption{Heatmap of languages in HLPSC according to endangerment status}\label{fig:heatmap}
\end{centering}
\end{figure*}

@@

Although data size matters in general NLP, \emph{universality} is the utmost priority for a universal corpus. We chose to focus on the following data sources, because they include a large number of languages, include several parallel texts, and demonstrate a variety of data types which a linguist might encounter (structured, semi-structured, unstructured):

\begin{enumerate}
\item The Online Database of Interlinear Text (ODIN)
\item The Omniglot website
\item The Universal Declaration of Human Rights (UDHR)
\item Wikipedia
\end{enumerate}

>>>>

\begin{figure*}[t]
\begin{centering}
\includegraphics[scale=0.63]{heatmap-numbers.png}
%\includegraphics{heatmap-numbers.png}
\caption{Heatmap of languages in SeedLing according to endangerment status}\label{fig:heatmap}
\end{centering}
\end{figure*}

@@

Although data size matters in general NLP, \emph{universality} is the utmost priority for a universal corpus. We chose to focus on the following data sources, because they include a large number of languages, include several parallel texts, and demonstrate a variety of data types which a linguist might encounter (structured, semi-structured, unstructured): The Online Database of Interlinear Text (ODIN), The Omniglot website, The Universal Declaration of Human Rights (UHDR), and Wikipedia.

% \begin{enumerate}
% \item The Online Database of Interlinear Text (ODIN)
% \item The Omniglot website
% \item The Universal Declaration of Human Rights (UDHR)
% \item Wikipedia
% \end{enumerate}

### 2014-03-20 docs/computEL/paper.tex
Major edits to intro
new heatmap file, added to data.tex
various small mods
added bib citations to cluster section

\usepackage[utf8]{inputenc}

@@

\title{Getting the Ball Rolling: Producing a Foundation Text for a Universal Corpus of the World's Languages}

@@

\begin{document}
\maketitle
\begin{abstract}
\textcolor{blue}{[revisit title, revisit abstract]}
The study of endangered languages is constrained by a lack of data. Existing corpora collections are limited in the range of languages covered, in standardisation, or in machine-readability. This makes the situation even worse for the computational linguist, especially one who would like to take a cross-linguistic or typological approach. We first survey existing efforts to compile cross-linguistic resources, then describe our own approach and give an example application - language clustering. To build the foundation text for a Universal Corpus, we crawled and cleaned texts from several web sources that contain data from a large number languages, and converted them into a standardised form consistent with the guidelines set out by \newcite{abney2010universal}. The resulting corpus is more easily-accessible and machine-readable than any of the underlying data sources, and represents a significant base corpus for researchers to draw on and add to in the future.
\end{abstract}

\input{intro}
\input{relwork}
%\input{liling}
\input{data}
\input{clean}
\input{cluster}
\input{corpus}
\input{concl}

>>>>

\usepackage[utf8]{inputenc}
\usepackage{graphicx}

@@

\title{SeedLing: Building and using a seed corpus \\ for the Human Language Project}
%Getting the Ball Rolling: Producing a Foundation Text for a Universal Corpus of the World's Languages}

@@

\begin{document}
\maketitle
\begin{abstract}
\textcolor{blue}{[revisit abstract]}
The study of endangered languages is constrained by a lack of data. Existing corpora collections are limited in the range of languages covered, in standardisation, or in machine-readability. This makes the situation even worse for the computational linguist, especially one who would like to take a cross-linguistic or typological approach. We first survey existing efforts to compile cross-linguistic resources, then describe our own approach and give an example application - language clustering. To build the foundation text for a Universal Corpus, we crawled and cleaned texts from several web sources that contain data from a large number languages, and converted them into a standardised form consistent with the guidelines set out by \newcite{abney2010universal}. The resulting corpus is more easily-accessible and machine-readable than any of the underlying data sources, and represents a significant base corpus for researchers to draw on and add to in the future.
\end{abstract}

\input{intro}
\input{relwork}
%\input{liling}
\input{data}
\input{clean}
\input{cluster}
%\input{corpus}
\input{concl}
